troubleshooting

### 2024-09-26 README.Rmd
Passed Rcmd Check

ggplot2::theme_set(
  ggplot2::theme_minimal()
)

```

@@

[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test
coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=main)
[![BioC
status](http://www.bioconductor.org/shields/build/release/bioc/mspms.svg)](https://bioconductor.org/checkResults/release/bioc-LATEST/mspms)

@@

The goal of mspms is provide a R Package that can be used to analyze data from
the [Multiplex Substrate Profiling by Mass Spectrometry (MSP-MS)
method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

@@

@@

You can install the development version of mspms from github.

```{r,eval = FALSE}
devtools::install_github("baynec2/mspms",
                         ref = "develop")
```

@@

```{r,eval = FALSE}
mspms::generate_report(
  prepared_data = mspms::peaks_prepared_data,
  colData = mspms::design_matrix,
  outdir = "../Desktop/mspms_report"
)
```

@@

#### Uploading your colData

@@

```{r}
colData = readr::read_csv("inst/extdata/colData.csv")
usethis::use_data(colData,overwrite = TRUE)

@@

# File path of peaks lfq file
lfq_filename <- "inst/extdata/peaks_protein-peptides-lfq.csv"

# Prepare the data 
peaks_prepared_data <- mspms::prepare_peaks(lfq_filename,
                                            colData,
                                            0.3,
                                            n_residues = 4)

usethis::use_data(peaks_prepared_data,overwrite = TRUE)

@@

```{r}

```

@@

```{r}
lfq_file_path = "inst/extdata/fragpipe_combined_peptide.tsv"

fragpipe_prepared_data = prepare_fragpipe(lfq_file_path,
                                          colData)

@@

Data processing includes imputation and normalization of the proteomics data.

**Add more info**

```{r}
# does set seed matter?
set.seed(2)
processed_qf = process_qf(peaks_prepared_data)

usethis::use_data(processed_qf,overwrite = TRUE)

processed_qf
```

@@

```{r}
log2fc_t_test_data = mspms::log2fc_t_test(processed_qf = processed_qf)

usethis::use_data(log2fc_t_test_data,overwrite = TRUE)
```

@@

We provide a number of data visualizations as part of the mspms_package.

@@

```{r}
plot_qc_check(processed_qf,
              full_length_threshold = 10,
              cleavage_product_threshold = 5)
```

@@

@@

```{r}
plot_heatmap(processed_qf)
```

@@

```{r}
plot_pca(processed_qf)
```

@@

```{r}
sig_cleavage_data = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3)

p1 = mspms::plot_cleavages_per_pos(sig_cleavage_data )

@@

```{r}
catA_sig_cleavages = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
  dplyr::filter(condition == "CatA") %>% 
  dplyr::pull(cleavage_seq) %>% 
  unique()
```

@@

```{r}
all_possible_8mers_from_228_library = calculate_all_cleavages(mspms::peptide_library$library_real_sequence,n_AA_after_cleavage = 4)

usethis::use_data(all_possible_8mers_from_228_library,overwrite = TRUE)
```

@@

```{r}
plot_icelogo(catA_sig_cleavages,
             background_universe = all_possible_8mers_from_228_library)
```

@@

```{r,fig.height = 10, fig.width = 4}
sig_cleavage_data = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3)

@@

```{r}
max_log2fc_pep = log2fc_t_test_data %>% 
  dplyr::filter(p.adj <= 0.05,log2fc > 3) %>% 
  dplyr::filter(log2fc ==max(log2fc)) %>% 
  pull(peptide)

```

We can then filter our QFeatures object to only include this peptide

```{r}
filter = QFeatures::VariableFilter(field = "peptide",
               value = max_log2fc_pep,
               condition = "==")

sel = QFeatures::filterFeatures(processed_qf,filter)
```

And then plot as follows:

```{r}
plot_time_course(sel,summarize_by_mean = FALSE)
```

>>>>

ggplot2::theme_set(
  ggplot2::theme_minimal()
)
```

<img src="man/figures/logo.png" align="right" width="170"/>

@@

[![R-CMD-check](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/baynec2/mspms/actions/workflows/R-CMD-check.yaml)
[![Codecov test
coverage](https://codecov.io/gh/baynec2/mspms/branch/main/graph/badge.svg)](https://app.codecov.io/gh/baynec2/mspms?branch=bioconductor)

@@

## Introduction

The goal of mspms is provide a R Package that can be used to easily and
reproducibly analyze data resulting from the [Multiplex Substrate Profiling by
Mass Spectrometry (MSP-MS) method](https://pubmed.ncbi.nlm.nih.gov/36948708/).

@@

Multiplex Substrate Profiling by Mass Spectrometry
[(MSP-MS)](https://pubmed.ncbi.nlm.nih.gov/36948708/) is a powerful method used
to determine the substrate specificity of proteases. This method is of interest
for groups interested in the study of proteases and their role as regulators of
biological pathways, whether applied to the study of disease states, the
development of diagnostic and prognostic tests, generation of tool compounds, or
rational design of protease targeting therapeutics. Analysis of the MS based
data produced by MSP-MS is a multi-step process involving detection and
quantification of peptides, imputation, normalization, cleavage sequence
identification, statistics, and data visualizations. This process can be
challenging, especially for scientists with limited programming experience. To
overcome these issues, we developed the mspms R package to facilitate the
analysis of MSP-MS data utilizing good software/data analysis practices.

mspms differs from existing proteomics packages that are available in the
Bioconductor project in that it is specifically designed to analyze MSP-MS data.
This involves unique preprocessing steps and data visualizations tailored to
this specific assay

In order to do so, mspms uses several excellent packages from the Bioconductor
project to provide a framework to easily and reproducibly analyze MSP-MS data.
These include:

-   QFeatures: management and processing of peptide quantitative features and
    sample metadata.

-   MsCoreUtils: log2 transformation, imputation (QRILC), and normalization
    (center.median).

Some excellent non-Bioconductor packages are also utilized.

-   ggplot2: data visualizations

-   heatmaply: interactive heatmaps

-   rstatix: conduct statistics

-   ggseqlogo: iceLogo motif visualization.

-   downloadthis: gives users the ability to download data from standard .html
    report generated by mspms::generate_report().

Lastly, mspms takes advantage of iceLogos ([Colaert, N. et al. Nature Methods 6,
786-787 (2009)](http://www.ncbi.nlm.nih.gov/pubmed/19876014)) to viusalize
overrepresented amino acid motifs relative to a background set by implementing
components of the Java software in R.

@@

You can install the stable release from Bioconductor:

Alternatively, you can install the development version from github.

```{r,eval = FALSE}
devtools::install_github("baynec2/mspms",
  ref = "develop"
)
```

@@

```{r,eval = FALSE}
mspms::generate_report(
  prepared_data = mspms::peaks_prepared_data,
  outdir = "../Desktop/mspms_report"
)
```

@@

#### A note on colData

@@

```{r}
colData <- readr::read_csv(system.file("extdata/colData.csv", package = "mspms"))

@@

# File path of peaks lfq file
lfq_filepath <- system.file("extdata/peaks_protein-peptides-lfq.csv",
  package = "mspms"
)

colData_filepath <- system.file("extdata/colData.csv", package = "mspms")

# Prepare the data
peaks_prepared_data <- mspms::prepare_peaks(lfq_filepath,
  colData_filepath,
  0.3,
  n_residues = 4
)

@@

```{r}
```

@@

```{r}
combined_peptide_filepath <- system.file("extdata/fragpipe_combined_peptide.tsv",
  package = "mspms"
)

fragpipe_prepared_data <- prepare_fragpipe(
  combined_peptide_filepath,
  colData_filepath
)

@@

Data processing includes imputation and normalization of the proteomics data.
This is all done in a QFeatures object using MSCoreUtils methods. Data from 
each step is stored as a SummarizedExperiment within the QFeatures object with
the indicated name.

1. Data is first log2 transformed ("peptides_log").   
2. Data is normalized using the center.median method ("peptides_log_norm").  
3. Missing values are then imputed using the QRILC method 
("peptides_log_impute_norm"). 
4. Data is then reverse log2 transformed ("peptides_norm").

```{r}
processed_qf <- process_qf(peaks_prepared_data)
processed_qf
```

@@

```{r}
log2fc_t_test_data <- mspms::log2fc_t_test(processed_qf = processed_qf)
```

@@

We provide a number of data visualizations as part of the mspms package.

@@

```{r}
plot_qc_check(processed_qf,
  full_length_threshold = 10,
  cleavage_product_threshold = 5
)
```

@@

### Tidying our data

We can then convert our QFeatures object into a "long" tibble using the 
mspms_tidy function.

This makes the data easy to manipulate prior to ggplot2 or plotly based plotting 
functions.

```{r}
mspms_tidy_data <- mspms_tidy(processed_qf)
```

@@

```{r}
plot_heatmap(mspms_tidy_data)
```

@@

```{r}
plot_pca(mspms_tidy_data, value_colname = "peptides_norm")
```

@@

```{r}
sig_cleavage_data <- log2fc_t_test_data %>%
  dplyr::filter(p.adj <= 0.05, log2fc > 3)

p1 <- mspms::plot_cleavages_per_pos(sig_cleavage_data)

@@

```{r}
catA_sig_cleavages <- log2fc_t_test_data %>%
  dplyr::filter(p.adj <= 0.05, log2fc > 3) %>%
  dplyr::filter(condition == "CatA") %>%
  dplyr::pull(cleavage_seq) %>%
  unique()
```

@@

```{r}
all_possible_8mers_from_228_library <- calculate_all_cleavages(
  mspms::peptide_library$library_real_sequence,
  n_AA_after_cleavage = 4
)
```

@@

```{r}
plot_icelogo(catA_sig_cleavages,
  background_universe = all_possible_8mers_from_228_library
)
```

@@

```{r,fig.height = 10, fig.width = 4}
sig_cleavage_data <- log2fc_t_test_data %>%
  dplyr::filter(p.adj <= 0.05, log2fc > 3)

@@

```{r}
max_log2fc_pep <- log2fc_t_test_data %>%
  dplyr::filter(p.adj <= 0.05, log2fc > 3) %>%
  dplyr::filter(log2fc == max(log2fc)) %>%
  pull(peptide)
```

We can then filter our tidy mspms data to only include this peptide and plot

```{r}
p1 <- mspms_tidy_data %>%
  dplyr::filter(peptide == max_log2fc_pep) %>%
  plot_time_course()

```{r}
sessionInfo()
```
